---
title: "Application-Difference between Lasomo and Lomo"
output:
  pdf_document: default
  html_document: default
---

```{r include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
suppressPackageStartupMessages(library(tidyverse))
library(dplyr)
library(magrittr)
library(lubridate)
library(pander)
library(ggpubr) # ggarrange
library(GGally) 
suppressPackageStartupMessages(library(GGally)) # for 'ggpairs'
library(here)
setwd(here())
```
```{r}
source("../Rcode/relative_wis.R") 
```

# Find a model that started submitting forecasts after November 2020

```{r}
# forecasts <- readRDS("../Data/forecast_data_aligned_death_10models_june20-nov22.rds")
# 
# # all models in the 'forecasts' data set
# models <- forecasts$model %>% unique()
# 
# # models before November 2020
# models_before_nov2020 <- forecasts %>% 
#         filter(forecast_date < "2020-11-01") %>% 
#         select(model) %>% pull() %>% unique()
# 
# # model that submitted forecasts after November 1st, 2020
# setdiff(models, models_before_nov2020 )
# # [1] "USC-SI_kJalpha" 
# 
# # First forecast date of the model "USC-SI_kJalpha"
# forecasts %>% 
#         filter(model == "USC-SI_kJalpha") %>% 
#         select(forecast_date) %>% pull() %>% unique() %>% min()
# # [1] "2020-11-15"
```

# Submission Proportion

```{r submission-proportion-wis, include=FALSE}
df_all_scores_extended <- readRDS("../Data/df_all_scores_extended.rds")
props <- df_all_scores_extended %>%
        group_by(model) %>%
        summarise(#na_prop = sum(is.na(wis))/21800,
                  submission_proportion = round(sum(!is.na(wis))/21800 * 100, 1),
                  submission_count = sum(!is.na(wis))) %>%
        ungroup() %>%
        arrange(desc(submission_proportion)) %>% 
        rename(Model = model,
                 "Submission rate (%)" = submission_proportion,
                 "Number of predictions (total:21800)" = submission_count)
```


# Compare scores when NA values are replaced by the worst score

For each model, location (l), time (t), horizon (h), replace NAs by the worst scores; 
by max WIS over all non-NA models for WIS, by 0 for rank, and by min importance scores over all non-NA models.

```{r table, include=FALSE}
df_all_scores_extended_NAworst <- readRDS("../Data/df_all_scores_extended_NAworst.rds") 

# compute relative wis compared to the "COVIDhub-baseline" model
relWIS <- relative_wis(df_all_scores_extended_NAworst)

compare_NAworst <- df_all_scores_extended_NAworst %>% 
        group_by(model) %>% 
        summarise(meanWIS = mean(wis),
                  meanRank = mean(std_rank),
                  meanImpScore_shapley_eq = mean(ImpScore_shapley_eq),
                  meanImpScore_shapley_wt = mean(ImpScore_shapley_wt),
                  meanImpScore_lomo_eq = mean(ImpScore_lomo_eq),
                  meanImpScore_lomo_wt = mean(ImpScore_lomo_wt)) %>% 
        ungroup() %>% 
        left_join(relWIS, by="model") %>% 
        mutate(model = reorder(factor(model), meanWIS)) %>%
        arrange(meanWIS) %>% 
        relocate(relWIS, .after = meanWIS) %>% 
        mutate_if(is.numeric, ~round(., 2))
# saveRDS(compare_NAworst, "../Data/compare_NAworst.rds")
compare_NAworst <- readRDS("../Data/compare_NAworst.rds")
compare_NAworst %>% 
        rename(WIS=meanWIS, 
               Rank=meanRank, 
               Shapley.eq = meanImpScore_shapley_eq,
               Shapley.wt = meanImpScore_shapley_wt,
               LOMO.eq = meanImpScore_lomo_eq,
               LOMO.wt = meanImpScore_lomo_wt) %>% 
        pander(caption = "Summary of mean WIS, rank, importance scores 
               with untrained and trained ensembles 
               by replaceing NA values with the worst score.",
               justify = "right", 
               split.table = Inf)
```


```{r plots, include=FALSE}
g1 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanWIS), color = "deeppink") 
g2 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanRank), color = "green3") +
        scale_x_discrete(guide = guide_axis(angle = 30)) +
        labs(x=NULL)
g3 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanImpScore_shapley_eq), color = "red") +
        labs(y="Shapley.eq") 
g4 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanImpScore_lomo_eq), color = "blue") +
        labs(y="LOMO.eq") +
        scale_x_discrete(guide = guide_axis(angle = 30))

g5 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanImpScore_shapley_wt), color = "red2") +
        labs(y="Shapley.wt") 
g6 <- compare_NAworst %>% ggplot() + 
        geom_point(aes(model, meanImpScore_lomo_wt), color = "blue2") +
        labs(y="LOMO.wt") +
        scale_x_discrete(guide = guide_axis(angle = 30))


remove_x <- theme(
  axis.text.x = element_blank(),
  axis.ticks.x = element_blank(),
  axis.title.x = element_blank()
)

library(ggpubr) #annotate_figure

plots_wis_rk <- ggarrange(g1 + remove_x, g2,ncol = 1, heights = c(0.65, 1))  
annotate_figure(plots_wis_rk, top = text_grob("mean WIS and mean Rank", 
               face = "bold", size = 14))

plots_untrained <- ggarrange(g3 + remove_x, g4,ncol = 1, heights = c(0.65, 1)) 
annotate_figure(plots_untrained, top = text_grob("mean Importance Score (untrained) ", 
               face = "bold", size = 14))

plots_trained <- ggarrange(g5 + remove_x, g6,ncol = 1, heights = c(0.65, 1)) 
annotate_figure(plots_trained, top = text_grob("mean Importance Score (trained) ", 
               face = "bold", size = 14))
```


# Summary of relationships between metrics 

```{r pairs-plot-NAworst, message=FALSE}
compare_NAworst <- readRDS("../Data/compare_NAworst.rds")
d_NAworst <- compare_NAworst %>% 
        mutate(neg_WIS=-meanWIS) %>% 
        relocate(neg_WIS, .after=model) 

colnames <- c(paste0("-WIS"), #paste0("Std.rank"), 
              "Phi[eq]^{lasomo}", "Phi[wt]^{lasomo}", 
              "Phi[eq]^{lomo}", "Phi[wt]^{lomo}")

p.pairs <- ggpairs(
        d_NAworst %>% select(c(2,6:9)),
        columnLabels = colnames,
        labeller = "label_parsed"
        ) + 
        theme(axis.text.x=element_text(angle = 45, vjust = 0.75, hjust=0.75),
              strip.text = element_text(size = 14),
              panel.spacing = unit(0.65, "lines"))

p.pairs 

pdf(file="../Plot/plot-corr-NAworst.pdf",
    width = 10,
    height = 5)
p.pairs
dev.off()
```

# Plot to illustrate the rank changes

```{r}
library(ggbump)

df_rank <- compare_NAworst %>%
        mutate(WIS_rank = rank(meanWIS), 
               lasomo_eq_rank = rank(-meanImpScore_shapley_eq),
               lasomo_wt_rank = rank(-meanImpScore_shapley_wt),
               lomo_eq_rank = rank(-meanImpScore_lomo_eq),
               lomo_wt_rank = rank(-meanImpScore_lomo_wt)) %>% 
        select(model, WIS_rank, lasomo_eq_rank, lasomo_wt_rank, lomo_eq_rank, lomo_wt_rank) 

# Plot: WIS vs. Phi[eq]^{lasomo}
rank_lasomo_eq <- df_rank %>% 
        mutate(line_type = case_when(
                                WIS_rank < lasomo_eq_rank ~ "twodash",
                                WIS_rank > lasomo_eq_rank ~ "dotted",
                                TRUE ~ "solid"
                                )
               ) %>%
        select(model, WIS_rank, lasomo_eq_rank, line_type) %>%
        pivot_longer(cols = c(WIS_rank, lasomo_eq_rank), 
                     names_to = "Metric", 
                     values_to = "Rank") %>% 
        ggplot(aes(x = Metric, y = Rank, group = model, linetype = line_type)) +
        geom_bump(linewidth=2) +
        geom_point(size=0, alpha = 0.1) +
        scale_x_discrete(limits = rev, labels = c("WIS", expression(Phi[eq]^{lasomo})),
                         expand = c(0.45, 0.45)) +
        scale_y_continuous(trans="reverse", breaks = 1:10) +
        labs(title = bquote(bold("(a) WIS vs. " * Phi[eq]^{lasomo})),
             x = "") +
        # Add labels for each line
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "lasomo_eq_rank"),  # Place label at the last point
                  hjust = "left",  
                  nudge_x = 0.03,
                  size = 6 )   +
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "WIS_rank"),  # Place label at the starting point
                  hjust = "right", 
                  nudge_x = -0.03,
                  size = 6)   +
        theme_bw() +
        theme(legend.position = "none",
              axis.title = element_text(size = 22),
              axis.text.x = element_text(size = 22),
              axis.text.y = element_text(size = 18),
              plot.title = element_text(size=23)) 
 
# Plot: WIS vs. Phi[wt]^{lasomo}
rank_lasomo_wt <- df_rank %>% 
        mutate(line_type = case_when(
                                WIS_rank < lasomo_wt_rank ~ "twodash",
                                WIS_rank > lasomo_wt_rank ~ "dotted",
                                TRUE ~ "solid"
                                )
               ) %>%
        select(model, WIS_rank, lasomo_wt_rank, line_type) %>%
        pivot_longer(cols = c(WIS_rank, lasomo_wt_rank), 
                     names_to = "Metric", 
                     values_to = "Rank") %>% 
        ggplot(aes(x = Metric, y = Rank, group = model, linetype = line_type)) +
        geom_bump(linewidth=2) +
        geom_point(size=0, alpha = 0.1) +
        scale_x_discrete(limits = rev, labels = c("WIS", expression(Phi[wt]^{lasomo})),
                         expand = c(0.45, 0.45)) +
        scale_y_continuous(trans="reverse", breaks = 1:10) +
        labs(title = bquote(bold("(b) WIS vs. " * Phi[wt]^{lasomo})),
             x = "", y="") +
        # Add labels for each line
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "lasomo_wt_rank"),  # Place label at the last point
                  hjust = "left",  
                  nudge_x = 0.03,
                  size = 6 )   +
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "WIS_rank"),  # Place label at the starting point
                  hjust = "right", 
                  nudge_x = -0.03,
                  size = 6)   +
        theme_bw() +
        theme(legend.position = "none",
              axis.title = element_text(size = 22),
              axis.text.x = element_text(size = 22),
              axis.text.y = element_text(size = 18),
              plot.title = element_text(size=23))       

# Plot: WIS vs. Phi[eq]^{lomo}
rank_lomo_eq <- df_rank %>% 
        mutate(line_type = case_when(
                                WIS_rank < lomo_eq_rank ~ "twodash",
                                WIS_rank > lomo_eq_rank ~ "dotted",
                                TRUE ~ "solid"
                                )
               ) %>%
        select(model, WIS_rank, lomo_eq_rank, line_type) %>%
        pivot_longer(cols = c(WIS_rank, lomo_eq_rank), 
                     names_to = "Metric", 
                     values_to = "Rank") %>% 
        ggplot(aes(x = Metric, y = Rank, group = model, linetype = line_type)) +
        geom_bump(linewidth=2) +
        geom_point(size=0, alpha = 0.1) +
        scale_x_discrete(limits = rev, labels = c("WIS", expression(Phi[eq]^{lomo})),
                         expand = c(0.45, 0.45)) +
        scale_y_continuous(trans="reverse", breaks = 1:10) +
        labs(title = bquote(bold("(c) WIS vs. " * Phi[eq]^{lomo}))) +
        # Add labels for each line
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "lomo_eq_rank"),  # Place label at the last point
                  hjust = "left",  
                  nudge_x = 0.03,
                  size = 6 )   +
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "WIS_rank"),  # Place label at the starting point
                  hjust = "right", 
                  nudge_x = -0.03,
                  size = 6)   +
        theme_bw() +
        theme(legend.position = "none",
              axis.title = element_text(size = 22),
              axis.text.x = element_text(size = 22),
              axis.text.y = element_text(size = 18),
              plot.title = element_text(size=23)) 

# Plot: WIS vs. Phi[wt]^{lomo}
rank_lomo_wt <- df_rank %>% 
        mutate(line_type = case_when(
                                WIS_rank < lomo_wt_rank ~ "twodash",
                                WIS_rank > lomo_wt_rank ~ "dotted",
                                TRUE ~ "solid"
                                )
               ) %>%
        select(model, WIS_rank, lomo_wt_rank, line_type) %>%
        pivot_longer(cols = c(WIS_rank, lomo_wt_rank), 
                     names_to = "Metric", 
                     values_to = "Rank") %>% 
        ggplot(aes(x = Metric, y = Rank, group = model, linetype = line_type)) +
        geom_bump(linewidth=2) +
        geom_point(size=0, alpha = 0.1) +
        scale_x_discrete(limits = rev, labels = c("WIS", expression(Phi[wt]^{lomo})),
                         expand = c(0.45, 0.45)) +
        scale_y_continuous(trans="reverse", breaks = 1:10) +
        labs(title = bquote(bold("(d) WIS vs. " * Phi[wt]^{lomo})), y=" ") +
        # Add labels for each line
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "lomo_wt_rank"),  # Place label at the last point
                  hjust = "left",  
                  nudge_x = 0.03,
                  size = 6 )   +
        geom_text(aes(label = model), 
                  data = . %>% filter(Metric == "WIS_rank"),  # Place label at the starting point
                  hjust = "right", 
                  nudge_x = -0.03,
                  size = 6)   +
        theme_bw() +
        theme(legend.position = "none",
              axis.title = element_text(size = 22),
              axis.text.x = element_text(size = 22),
              axis.text.y = element_text(size = 18),
              plot.title = element_text(size=23)) 
```

```{r}
pdf(file="../Plot/rank-changes-lasomo-eq-NAworst.pdf",
    width = 10,
    height = 8)
rank_lasomo_eq
dev.off()

pdf(file="../Plot/rank-changes-lasomo-wt-NAworst.pdf",
    width = 10,
    height = 8)
rank_lasomo_wt
dev.off()

pdf(file="../Plot/rank-changes-lomo-eq-NAworst.pdf",
    width = 10,
    height = 8)
rank_lomo_eq
dev.off()

pdf(file="../Plot/rank-changes-lomo-wt-NAworst.pdf",
    width = 10,
    height = 8)
rank_lomo_wt
dev.off()
```






