---
title: "Application-case_study"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(covidHubUtils)
library(dplyr)
library(lubridate)
library(stringr)
library(ggplot2)
library(ggrepel)
library(ggridges)
library(ggpubr)
library(gridExtra)
library(grid)
library(doParallel)
library(tidyverse)
theme_set(theme_bw())
library(pander)
library(DT)
library(ggalt)  #for geom_dumbbell()
library(pander)
library(kableExtra)
library(ggprism)
library(patchwork)
library(here)
setwd(here())
```

# Load forecast data of death in MA in 2021

This dataset contains 9 models, their importance score, and WIS on each target end date for 4-week ahead incident death.

```{r load_data_forecast_ma2021, echo = FALSE}
forecast_ma2021_full <- readRDS("../Data/forecast_death_ma2021.rds")

forecast_ma2021 <- forecast_ma2021_full %>% 
  rename(f.median = 'X0.5') %>% 
  select(model, importance, wis, target_end_date, location, target, 
         dispersion, overpred, underpred, f.median) 
```


# Scatterplots of importance vs. WIS by model

```{r plot_importance-vs-wis-by-model, fig.pos='h', fig.cap="left-aligned"}
dat <- forecast_ma2021 %>% 
  filter(!grepl('Ens', model)) %>% 
  mutate(target_end_date=str_remove(target_end_date, "2021-")) 
p <- dat %>% 
        ggplot(aes(x=wis, y= importance)) +
        geom_hline(yintercept = 0, linetype="dashed", 
                      size=0.5, colour="black") +
        geom_point(shape=24, color = "grey57", size=2) +
        geom_point(data = dat %>% filter(target_end_date == "12-25"), 
                   shape=21, fill = "black", color="black", size=2.5)  +
        facet_wrap(~ model) +
        labs(#title = "Model importance vs. WIS by Model",
                x = "WIS",
                y = "Importance") + 
        theme(axis.text=element_text(size=10),
              axis.title=element_text(size=16),
              legend.position = "none",
              strip.text = element_text(size = 10))

pdf(file="../Plot/scatterplot-wis_imp2021.pdf",
    width = 10,
    height = 5)
p
dev.off()
```


# Plot of Importance vs. WIS on 2021-12-25 
 
This plot shows that there is an inaccurate but important model because, while this model had a large positive bias, it was the only model that was biased in that direction on this particular week.
That bias, in this situation, therefore made this model an “important” counter-weight to all of the other models.

```{r}
dat <- forecast_ma2021 %>% 
        filter(!grepl('Ens', model)) %>% 
        filter(target_end_date == "2021-12-25")
p.imp_wis.1225 <- dat %>% 
        ggplot(aes(x=wis, y = importance)) + 
        geom_point(shape=21, fill = "black", color="black", size=3) + 
        geom_text_repel(data=subset(dat, model == "CovidAnalytics-DELPHI"),
                         aes(label = model),
                        size = 5,
                         nudge_x = -32,
                         nudge_y = -2,
                         box.padding = 0.5,
                         point.padding = 0.5,
                         arrow = arrow(length = unit(0.03, "npc"))) + 
        xlim(0,130) + 
        labs(y = "Importance",
             x = "WIS") + 
        theme(axis.text=element_text(size=14),
              axis.title=element_text(size=16),
              legend.position="none")


pdf(file="../Plot/plot-wis_imp_20211225.pdf",
    width = 10,
    height = 5)
p.imp_wis.1225 
dev.off()
```


# Plot of prediction intervals of 9 models and ensembles constructed by leave-one-model-out algorithm

```{r}
truth.ma.2021 <- read_csv("../Data/truth_MA_2021.csv") %>% 
  select(target_end_date, value) %>% 
  rename(truth = value)

y.val= truth.ma.2021$truth[truth.ma.2021$target_end_date == "2021-12-25"] 

# lock in factor level order
forecast_ma2021_full.v2 <- forecast_ma2021_full %>% 
  mutate(model_order = factor(model, levels = unique(model))) %>% 
  relocate(model_order)

d <- forecast_ma2021_full.v2 %>% 
        filter(target_end_date == "2021-12-25") %>% 
        select(model_order, 8:30) 
model_sorted <- dat %>% arrange(desc(importance)) %>% select(model)
d1 <- d[c(10, 1:9), ]
d2 <- d[c(10:19), ]

# Prediction intervals of single models
p1 <- d[c(1:9), ] %>%  
        ggplot(aes(y = factor(model_order, levels = model_sorted$model))) + 
        geom_dumbbell(aes(x = X0.025, xend = X0.975)) +
        geom_point(aes(x = X0.5 ) ) + 
        scale_y_discrete(labels = function(x, ...) gsub('-', '\n-', x)) +
        geom_vline(xintercept = y.val, linetype="dashed", 
                      size=0.5, colour="darkred") +
        labs(x="Value",
             title="95% Prediction Intervals of Individual Forecasts", 
             y="Model") + 
        theme(axis.text.x = element_text(size=8, angle = 30, hjust = 0.85),
              legend.position = "none") +
        theme(axis.title.x = element_text(size = 14),
              axis.text.x = element_text(size = 8),
              axis.title.y = element_text(size = 14)) +
        theme(text = element_text(size = 10)) + coord_flip() 

```

```{r}
pdf(file="../Plot/95PI1.pdf",
    width = 6,
    height = 3)
p1
dev.off()
```

```{r}
# Prediction intervals of ensemble models

d2.copy <- d2 %>% mutate(model_renamed = model_order) %>% relocate(model_renamed)
d2.copy$model_renamed <- str_remove(d2.copy$model_renamed, pattern = "Ens.wo.")
d2.copy <- d2.copy %>% 
        mutate( model_renamed = str_replace(model_renamed, 
                                            "Ensemble.all", 
                                            "None-(ensemble of all)"))
extended_model_sorted <- c(model_sorted$model, "None-(ensemble of all)")
breaks2 <- c(seq(0, 700, by=100))
labels2 <- as.character(breaks2)

p2 <- d2.copy %>%
        ggplot(aes(y = factor(model_renamed, levels = extended_model_sorted))) +
        geom_dumbbell(aes(x = X0.025, xend = X0.975)) +
        geom_point(aes(x = X0.5 ) ) +
        scale_x_continuous(limits = c(-10, 500), breaks = breaks2, labels = labels2) +
        scale_y_discrete(labels = function(x, ...) gsub('-', '\n-', x)) +
        geom_vline(xintercept = y.val, linetype="dashed", 
                      size=0.5, colour="darkred") +
        labs(x="Value", y="Left-out model", 
             title="95% Prediction Intervals of Ensembles Built Leaving One Model Out") + 
        theme(axis.text.x = element_text(size=8, angle = 30, hjust = 0.85),
              legend.position = "none") +
        theme(axis.title.x = element_text(size = 16),
              axis.text.x = element_text(size = 10),
              axis.title.y = element_text(size = 16)) +
        theme(text = element_text(size = 12))  + coord_flip()   
```


```{r}
pdf(file="../Plot/95PI2.pdf",
    width = 8,
    height = 4)
p2
dev.off()
```











