---
title: "Data Processing - Combine all score data into a single dataframe"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
# library(covidData)
library(scoringutils)
library(covidHubUtils)
library(hubEnsembles)
suppressPackageStartupMessages(library(tidyverse))
library(dplyr)
library(magrittr)
library(lubridate)
library(furrr)
library(purrr)
library(pander)
library(here)
setwd(here())
```


```{r read-forecasts-truth, include=FALSE}
forecast_data <- readRDS("../Data/forecast_data_aligned_death_10models_nov20-nov22.rds") %>%  
        select(-c(2,6,7)) %>%
        rename(forecast_date = reference_date,
               target_type = target_variable,
               prediction = value) %>% # Change the column names so that they meet the requirement for 'score()' in scoringutils package
        arrange(forecast_date, model)

truth_data <- readRDS("../Data/truth_data_upto2022.rds") # %>% select("location", "target_end_date", "value")

truth <- truth_data %>% select(location, target_end_date, value) %>% rename(true_value = value)
```

# Get WIS 

```{r calc-wis, include=FALSE}
# forecast_data <- readRDS("../Data/forecast_data_aligned_death_10models_nov20-nov22.rds") %>% select(-c(2)) %>% 
#         rename(forecast_date = reference_date) %>% # Change the column names  
#         arrange(forecast_date, model)
# truth_data <- readRDS("../Data/truth_data_upto2022.rds")
# 
# scores <- score_forecasts(
#   forecasts = forecast_data,
#   return_format = "wide",
#   metrics = "wis",
#   truth = truth_data,
#   use_median_as_point = TRUE
# )

# saveRDS(scores, "../Data/scores_wis.rds")
```


# Combine data frames of WIS and Importance scores with trained/untrained ensembles

```{r wis-shapley-both, include=FALSE}
# scores_wis <- readRDS("../Data/scores_wis.rds")
# wis <- scores_wis %>% select(model, location, horizon, forecast_date, wis)
shapley_trained <- readRDS("../Data/shapley_trained_completed.rds") %>%
        rename(ImpScore_shapley_wt = ImportanceScore) %>%
        transform(horizon=as.character(horizon),
                  forecast_date = as.Date(forecast_date))

shapley_untrained <- readRDS("../Data/shapley_untrained_completed.rds") %>%
        rename(ImpScore_shapley_eq = ImportanceScore) %>%
        transform(horizon=as.character(horizon),
                  forecast_date = as.Date(forecast_date))

lomo_untrained <- readRDS("../Data/lomo_untrained_completed.rds") %>%
        rename(ImpScore_lomo_eq = ImportanceScore) %>%
        transform(horizon=as.character(horizon),
                  forecast_date = as.Date(forecast_date))

lomo_trained <- readRDS("../Data/lomo_trained_completed.rds") %>%
        rename(ImpScore_lomo_wt = ImportanceScore) %>%
        transform(horizon=as.character(horizon),
                  forecast_date = as.Date(forecast_date))
```


```{r combine-wis-shapleys, include=FALSE}
# df_all_scores <- shapley_untrained %>%
#         left_join(shapley_trained, by = c("model", "horizon", "forecast_date", "location")) %>%
#         left_join(wis, by = c("model", "horizon", "forecast_date", "location")) %>%
#         left_join(lomo_untrained, by = c("model", "horizon", "forecast_date", "location")) %>%
#         left_join(lomo_trained, by = c("model", "horizon", "forecast_date", "location"))
```


# Get rank

```{r extend-data-to-include-std_rank, include=FALSE}
# # Add columns to calculate standardized WIS ranks
# num_mod_col <- df_all_scores %>%
#         dplyr::group_by(horizon, forecast_date, location) %>%
#         summarise(n_mod = sum(!is.na(wis)) , .groups = "drop")
# 
# rank_col <- df_all_scores %>% select(-model) %>%
#         dplyr::group_by(horizon, forecast_date, location) %>%
#         mutate(rank = rank(wis,na.last = "keep")) %>%
#         ungroup()
# 
# df_all_scores_extended <- df_all_scores %>%
#         left_join(num_mod_col,
#                   by = c("horizon", "forecast_date", "location")) %>%
#         left_join(rank_col,
#                   by = c("horizon", "forecast_date", "location", "wis",
#                          "ImpScore_shapley_wt", "ImpScore_shapley_eq",
#                          "ImpScore_lomo_wt","ImpScore_lomo_eq")) %>%
#         distinct() %>%
#         mutate(std_rank = 1-(rank-1)/(n_mod-1))
# 
# saveRDS(df_all_scores_extended, "../Data/df_all_scores_extended.rds")
```

```{r completeness-proportion, include=FALSE}
# (df_all_scores_extended %>% complete.cases() %>% sum()) / 218000

# => 0.9522752
```



```{r NA-proportion-all-metrics, include=FALSE}
# df_all_scores_extended %>% 
#         group_by(model) %>% 
#         summarise(na_prop_wis = sum(is.na(wis))/21800,
#                   na_prop_shapley_eq = sum(is.na(ImpScore_shapley_eq))/21800,
#                   na_prop_shapley_wt = sum(is.na(ImpScore_shapley_wt))/21800,
#                   na_prop_lomo_eq = sum(is.na(ImpScore_lomo_eq))/21800,
#                   na_prop_lomo_wt = sum(is.na(ImpScore_lomo_wt))/21800) %>% 
#         ungroup()

# => na proportion of USC-SI_kJalpha are different when using equally weighted ensemble from when using trained ensemble because we don't include a model that is in the current forecasts set but not in the training set. 
```



# 1. Replace NAs with the worst score

For each model, location (l), time (t), horizon (h), replace NAs 

- WIS : by max WIS over all non-NA models

- Rank : with zero

- Importance: by min importance over all non-NA models


```{r}
df_all_scores_extended <- readRDS("../Data/df_all_scores_extended.rds") 

df_all_scores_extended_NAworst <- df_all_scores_extended %>% #select(-model) %>%
        dplyr::group_by(horizon, forecast_date, location) %>%
        mutate_at( vars(wis), ~replace_na(., max(., na.rm=TRUE)) ) %>%
        mutate_at( vars(std_rank), ~replace_na(., 0) ) %>%
        mutate_at( vars(ImpScore_shapley_wt), ~replace_na(., min(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_shapley_eq), ~replace_na(., min(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_lomo_wt), ~replace_na(., min(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_lomo_eq), ~replace_na(., min(., na.rm=TRUE)) ) %>%
        ungroup()

saveRDS(df_all_scores_extended_NAworst, "../Data/df_all_scores_extended_NAworst.rds")
```


# 2. Replace NAs with the average score

For each model, location (l), time (t), horizon (h), replace NAs 

- WIS : by mean WIS over all non-NA models

- Rank : with 0.5

- Importance: by mean importance over all non-NA models


```{r}
df_all_scores_extended <- readRDS("../Data/df_all_scores_extended.rds")

df_all_scores_extended_NAavg <- df_all_scores_extended %>% #select(-model) %>%
        dplyr::group_by(horizon, forecast_date, location) %>%
        mutate_at( vars(wis), ~replace_na(., mean(., na.rm=TRUE)) ) %>%
        mutate_at( vars(std_rank), ~replace_na(., 0.5) ) %>%
        mutate_at( vars(ImpScore_shapley_wt), ~replace_na(., mean(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_shapley_eq), ~replace_na(., mean(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_lomo_wt), ~replace_na(., mean(., na.rm=TRUE)) ) %>%
        mutate_at( vars(ImpScore_lomo_eq), ~replace_na(., mean(., na.rm=TRUE)) ) %>%
        ungroup()

saveRDS(df_all_scores_extended_NAavg, "../Data/df_all_scores_extended_NAavg.rds")
```


# 3. Drop NAs

Drop all NAs

```{r}
df_all_scores_extended <- readRDS("../Data/df_all_scores_extended.rds")

df_all_scores_extended_NAdrop <- df_all_scores_extended %>% drop_na()
saveRDS(df_all_scores_extended_NAdrop, "../Data/df_all_scores_extended_NAdrop.rds")
```


